---
title: "groundwater_pollution_dk"
author: "Johan Horsmans"
date: "4/26/2021"
output: github_document
---
# Loading- and preprocessing data:

### Loading packages
```{r}
library(pacman)
p_load(sf, raster, dplyr, tmap, ggplot2, tidyverse, lubridate, sp, gstat, ggthemes, lmtest)
```

### Loading nitrate data:
```{r, results='hide'}
# Load the .csv-file as a dataframe and save it as "nitrate":
nitrate <- as.data.frame(read_csv("data/nitrate.csv")) 

# Rename columns to remove spaces and odd characters:
names(nitrate)[names(nitrate) == "Seneste"] <- "Seneste_m책ling" 
names(nitrate)[names(nitrate) == "Seneste mg/l"] <- "Seneste_mgl"
names(nitrate)[names(nitrate) == "Indtag topdybde"] <- "Indtag_topdybde" 
```

### Remove outliers:
```{r}
# Remove all nitrate-measures above 200:
nitrate <- nitrate %>%
  filter_at(vars("Seneste_mgl"), any_vars(. < 200))

# Remove all nitrate-measures below 0:
nitrate <- nitrate %>%
  filter_at(vars("Indtag_topdybde"), any_vars(. > 0))
```


### Processing the nitrate-dataframe to make it compatible with shapefile-format:
```{r}
# Transform coordinate column to characters:
nitrate$WKT <- as.character(nitrate$WKT)

# Use Regex to remove non-coordinate characters:
nitrate$WKT <- gsub("POINT \\(", "", nitrate$WKT)
nitrate$WKT <- gsub(")", "", nitrate$WKT)
```

### Separate into two columns:
```{r}
# Separate the longitude- and latitude coordinates into separate columns. The coordinates are separated by space (i.e. " "):
nitrate <- nitrate %>% 
  separate(col = WKT, into = c("longitude","latitude"), sep = " ") 
```

### Making it a shapefile:
```{r}
nitrate <- st_as_sf(nitrate, coords = c("longitude", "latitude"))
```

### Inspecting "nitrate":
```{r}
head(nitrate)
```

### Loading land-use data:
```{r}
# Loading "Markblok.shp" which contains all registered fields in Denmark:
all_fields <- st_read("data/Markblok.shp") %>% na.omit() 

# Convert "fields" to SpatialPolygons-object:
all_fields <- as_Spatial(all_fields$geometry)

# Load all registered ecological fields (from 2012 to 2020):
organic_fields_2012<-st_read("data/Oekologiske_arealer_2012.shp") %>% na.omit()
organic_fields_2013<-st_read("data/Oekologiske_arealer_2013.shp") %>% na.omit()
organic_fields_2014<-st_read("data/Oekologiske_arealer_2014.shp") %>% na.omit()
organic_fields_2015<-st_read("data/Oekologiske_arealer_2015.shp") %>% na.omit()
organic_fields_2016<-st_read("data/Oekologiske_arealer_2016.shp") %>% na.omit()
organic_fields_2017<-st_read("data/Oekologiske_arealer_2017.shp") %>% na.omit()
organic_fields_2018<-st_read("data/Oekologiske_arealer_2018.shp") %>% na.omit()
organic_fields_2019<-st_read("data/Oekologiske_arealer_2019.shp") %>% na.omit()
organic_fields_2020<-st_read("data/Oekologiske_arealer_2020.shp") %>% na.omit()

# Merge ecological field-geometries as "organic_fields":
organic_fields <- c(organic_fields_2012$geometry, organic_fields_2013$geometry, organic_fields_2014$geometry, organic_fields_2015$geometry, organic_fields_2016$geometry, organic_fields_2017$geometry, organic_fields_2018$geometry, organic_fields_2019$geometry, organic_fields_2020$geometry)

# Convert "organic_fields" to SpatialPolygons-object:
organic_fields <- as_Spatial(organic_fields)

# Make a new variable called "conventional_fields", where all polygons in "all_fields" that overlap with the polygons in "organic_fields" are removed:
conventional_fields <- all_fields[lengths(st_intersects(st_as_sf(all_fields),st_as_sf(organic_fields)))==0,]

# Assess how many organic fields have been filtered out:
length(all_fields) - length(conventional_fields)
```
### Inspecting "conventional_fields" and "organic_fields":
```{r}
# Inspect "conventional_fields" and "organic_fields":
print("conventional fields:")
head(conventional_fields)

print("organic fields:")
head(organic_fields)
```

### Plotting fields:
```{r}
# Plot organic fields:
plot(organic_fields, main = "Organic fields:")
```

```{r}
# Plot conventional fields
plot(conventional_fields, main = "Conventional fields:")
```

### Taking subset of conventional fields to reduce computational load (see written report for more info):
```{r}
# Subsetting "conventional_fields" to make it contain as many polygons as "organic_fields":
conventional_fields <- sample(conventional_fields, length(organic_fields))
```

### Setting CRS (for more info, see written report):
```{r}
# Set the projection of the nitrate data as EPSG 25832:
nitrate <- st_set_crs(nitrate, value = 25832)

# Set the projection of the organic fields as EPSG 25832:
proj4string(organic_fields) <- CRS("+init=epsg:25832")

# Set the projection of the conventional fields as EPSG 25832:
proj4string(conventional_fields) <- CRS("+init=epsg:25832")

# Transform the geometry of the data to the assigned CRS:
organic_fields <- spTransform(organic_fields, CRS("+init=epsg:25832"))
conventional_fields <- spTransform(conventional_fields, CRS("+init=epsg:25832"))
nitrate <- st_transform(nitrate, crs=25832)

# Verify the data has been assigned new CRS:
head(nitrate)
crs(organic_fields)
crs(conventional_fields)
```

### Filtering dates between 2012 and 2021:
```{r}
nitrate <- nitrate %>%
 select(Seneste_m책ling, Seneste_mgl, geometry, Indtag_topdybde) %>%
 filter(Seneste_m책ling >= as.Date("2012-01-01") & Seneste_m책ling <= as.Date("2021-03-10"))
```

### Remove duplicate entries:
```{r}
nitrate<-nitrate[!duplicated(nitrate$geometry), ]
```

### Plotting the nitrate-data:
```{r}
plot(st_geometry(nitrate), pch = 16, cex = 0.4, main = "Nitrate measurements:")
```

### Plotting points on top of organic fields (to assess CRS allignment):
```{r}
plot(st_geometry(nitrate), pch = 16, cex = 0.4, col = "red", main = "Nitrate measurements (red) and organic fields (black):")
plot(organic_fields, add = TRUE)
```

### Plotting points on top of all conventional fields (to assess CRS allignment)
```{r}
plot(st_geometry(nitrate), pch = 16, cex = 0.4, col = "red", main = "Nitrate measurements (red) and conventional fields (black):")
plot(conventional_fields, add = TRUE)
```

### Plotting points with nitrate per. mg/l metric:
```{r}
ggplot(nitrate, aes(colour = Seneste_mgl)) +
  geom_sf() + coord_sf(datum = st_crs(25832)) + theme_solarized() + labs(title = "Nitrate measurements:", color = "Nitrate (mg/L)") + scale_colour_gradientn(colors = c("darkgreen", "yellow", "red"))
```

# Exploratory statistics:

## The following sections is designed to create a plot with depth distributions for nitrate measurements carried out on conventional- and organic fields, respectively (see written report for more info).

### Creating a variable with convenional polygons in list-format:
```{r}
conventional_polygon_list <- lapply(conventional_fields@polygons, function(x) SpatialPolygons(list(x)))
```

### Creating separate columns for nitrate X- and Y coordinates:
```{r}
nitrate$X <- st_coordinates(nitrate)[,1]
nitrate$Y <- st_coordinates(nitrate)[,2]
```

### Finding nitrate measurements that fall within "conventional_fields" polygons:
```{r}
# Defining empty matrix for appending point-coordinates:
point <- matrix(ncol = 2)

# Create matrix with X- and Y coordinates for all nitrate measurements:
for (i in c(1:length(nitrate$geometry))){ # For the number of elements in in nitrate$geometry...
  p <- matrix(c(nitrate$X[i], nitrate$Y[i]), ncol = 2, byrow = TRUE) # ... iterate through X- and Y
  point <- rbind(p, point) # ... Append coordinates to "point"-matrix
}

# Remove NA in point matrix (generated when initializing the matrix):
point <- point %>% na.omit()

# Convert the point-matrix to "SpatialPoints"-object:
points <- SpatialPoints(point)

# Create variable called "conventional_overlap" with a function that goes through all points and polygons in the "conventional_polygon_list". If a point falls within a polygon it returns "1", if not it returns "NA":
conventional_overlap<-lapply(conventional_polygon_list, function(x) over(points, x))

# Create function that converts the "conventional_overlap"-variable to a regular list, rather than a nested list (for indexing purposes):
list_transform = function(x, sep = ".") {
    names(x) = paste0(seq_along(x))
    while(any(sapply(x, class) == "list")) {
        ind = sapply(x, class) == "list"
        temp = unlist(x[ind], recursive = FALSE)
        names(temp) = paste0(rep(names(x)[ind], lengths(x[ind])),
                             sep,
                             sequence(lengths(x[ind])))
        x = c(x[!ind], temp)
    }
    return(x)
}

# Apply the defined function to the "conventional_overlap"-variable:
conventional_overlap = list_transform(conventional_overlap)

# Create a variable "called "indexes", containing all nitrate-measurement-points falling within the conventional-field polygons. Note: the format is: "polygon:point":
indexes<-names(which(unlist(conventional_overlap) == 1))

# Print indexes:
indexes

# Remove the "polygon"-index (i.e. the index before the ".")
indexes <- gsub("^[^.]+.", "", indexes)

# Transform to numeric variable:
indexes<-as.numeric(indexes)

# Create a new column called "land_type" and fill it with "Out of bounds":
nitrate$land_type <- "Out of bounds"

# Change the value for "land_type" to "conventional" for all the nitrate measurements overlapping with conventional field polygons:
for (i in indexes) {
  nitrate$land_type[i] <- "conventional"
}
```

### Finding nitrate measurements that fall within "organic_fields" polygons:
```{r}
organic_polygon_list <- lapply(organic_fields@polygons, function(x) SpatialPolygons(list(x)))
```

```{r}
# Create variable called "organic_overlap" with a function that goes through all points and polygons in the "organic_polygon_list". If a point falls within a polygon it returns "1", if not it returns "NA":
organic_overlap<-lapply(organic_polygon_list, function(x) over(points, x))

# Apply the defined function to the "organic_overlap"-variable:
organic_overlap = list_transform(organic_overlap)

# Create a variable "called "indexes", containing all nitrate-measurement-points falling within the organic-field polygons. Note: the format is: "polygon:point":
indexes<-names(which(unlist(organic_overlap) == 1))

# Print indexes:
indexes

# Remove the "polygon"-index (i.e. the index before the ".")
indexes <- gsub("^[^.]+.", "", indexes)

# Transform to numeric variable:
indexes<-as.numeric(indexes)

# Change the value for "land_type" to "organic" for all the nitrate measurements overlapping with organic field polygons:
for (i in indexes) {
  nitrate$land_type[i] <- "organic"
}
```

### Plotting:
```{r}
ggplot() +
  geom_density(nitrate, mapping = aes(Indtag_topdybde, colour = land_type)) + theme_solarized() + scale_colour_solarized() + labs(title = "Depth density plot:", color = "Nitrate (mg/L)") + xlab("Measurement depth") + ylab("Density")
```


# Kriging:

### Make variogram by fitting X- and Y coordinates to nitrate mg/L:
```{r, out.width= "65%", out.extra='style="float:right; padding:10px"'}
vgm <- variogram(nitrate$Seneste_mgl ~ X + Y, nitrate)
plot(vgm)
```

As seen on the plot to the right, it appears that it is quite hard to find clear spatial correlations in the data. Nonetheless, it appears that measurements within 55000 meters of each other are more correlated than those further away. As such, we argue that this model can be used for kriging, since it only needs to interpolate data that is close to "real" data" (see written report for more info).

### Fit a model to the variogram:

```{r}
# Use the above variogram to eyeball model parameters:
nugget <- 550 # Initial value.
psill <- 70 # Sealing - nugget.
range <- 55000 # Point where spatial correlation stops. 

# Fit the variogram to data:
v_model <- fit.variogram(
  vgm, 
  model = vgm(
    model = "Ste",
    nugget = nugget,
    psill = psill,
    range = range,
    kappa = 0.5
  )
)

# Show the fitted variogram on top of the binned variogram (ADELA TEXT)
plot(vgm, model = v_model)
```

### Create kriging grid
```{r}
# Find the dimensions of the nitrate data:
st_bbox(nitrate)

# Using the st_bbox output, manually design a grid slightly larger than the bounding box (for prettier plots). Specify that the kriging resolution should be 2000 meter pr. estimate:
grid <- GridTopology(c(430734,6040448), c(2000, 2000), c(236, 190))

# Make the grid a GridTopology-object with the same CRS as the nitrate data:
gridpoints <- SpatialPoints(grid, proj4string = CRS(projection("+init=epsg:25832 +proj=utm +zone=32 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m
+no_defs")))

# Convert gridpoint to SpatialPixels:
spgrid <- SpatialPixels(gridpoints)

# Define coordinate names to be X and Y:
coordnames(spgrid) <- c("X", "Y")

# Plot the grid:
plot(spgrid)
```


### Use fitted variogram model to interpolate new data to the grid defined above


```{r}
# Force the nitrate CRS to be consistent with spgrid (otherwise error messages ensue):
nitrate_sp <- as(nitrate, "Spatial")
crs(nitrate_sp) <- crs(spgrid)

# Do kriging interpolations over the grid
nitrate_grid <- krige(Seneste_mgl ~ X + Y, nitrate_sp, newdata = spgrid, model = v_model)


# Plot the nitrate measurements on top of the kriged "nitrate_grid":

# Plotting raster:
tm_shape(nitrate_grid[1])  +
  tm_raster(title = "Nitrate mg/l", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_credits(text = "Johan Horsmans & Emil Jessen") +
  tm_layout(main.title = "Nitrate pollution map") +
  
# Nitrate overlayed on raster:
tm_shape(nitrate_grid[1])  +
  tm_raster(title = "Nitrate mg/l", 
            style = "cont",
            palette = "-RdYlGn") +
tm_shape(nitrate) +
  tm_dots()
```

### Load DK map
```{r}
DK <- st_read("data/denmark_administrative_outline_boundary.shp")
```

### Inspect CRS
```{r}
head(DK)
```

We see that the map has the wrong CRS (i.e. 4326).

### Set correct CRS:
```{r}
# Set the projection of the DK-map as EPSG 25832:
DK <- st_set_crs(DK, value = 4326)

# Transform the geometry of the map to the assigned CRS:
DK <- st_transform(DK, crs=25832)
```

### Inspect CRS:
```{r}
head(DK)
```
### Crop raster to fit DK map (i.e. remove excess interpolations):
```{r}
nitrate_raster <- raster(nitrate_grid)

nitrate_grid_cropped <- crop(nitrate_raster, extent(DK))
nitrate_grid_cropped <- raster::mask(nitrate_raster, DK)
```

### Make plot of interpolated data with DK map and nitrate measurements on top:
```{r}
# Plotting raster:
tm_shape(nitrate_grid_cropped)  +
  tm_raster(title = "Nitrate mg/l", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_credits(text = "Johan Horsmans & Emil Jessen") +
  tm_layout(main.title = "Nitrate pollution map",
            legend.position = c("right","top"),
            legend.bg.color = "white", legend.bg.alpha = .2, 
            legend.frame = "gray50",
            bg.color = "lightblue") +
  
# Points overlayed on raster:
tm_shape(nitrate) +
  tm_dots() +
  
# Adding DK-map:
tm_shape(DK) + 
  tm_polygons(alpha = 0.3)
```

# Statistical modeling:

### Extracting all kriged nitrate-values falling within organic fields:
```{r}
# Extract kriged nitrate data (r) that fall "within organic_fields":
organic_extraction <- suppressWarnings(raster::extract(x=nitrate_raster, y=organic_fields, fun=mean, df=TRUE, na.rm=TRUE))

# Add column land_type and specify that it should be "organic":
organic_extraction$land_type <- "organic"
```


### Extracting all kriged nitrate-values falling within conventional fields:

```{r}
# Extract kriged nitrate data (r) that fall "within conventional_fields":
conventional_extraction <- suppressWarnings(raster::extract(x=nitrate_raster, y=conventional_fields, fun=mean, df=TRUE, na.rm=TRUE))

# Add column land_type and specify that it should be "conventional":
conventional_extraction$land_type<-"conventional"
```

### Merge the extraced data:

```{r}
# Merge the extracted data as "merged":
merged <- rbind(organic_extraction, conventional_extraction)
```


Fit linear regression model with nitrate concentration predicted by land type:
```{r}
# Fit linear regression model with nitrate concentration (var1.pred) ~ land_type:
lm_model <- lm(var1.pred ~ land_type, data = merged)

# Perform coefficient t-test:
coeftest(lm_model, type = "HC1")
```

For interpretation of output, see written report.
